---
title: "Andrew's notes"
author: "Andrew MacDonald"
date: '2014-07-03'
output: 
  html_document: default
  md_document: default
---

# Notes on data structures

##### Quiz

Take this short quiz to determine if you need to read this chapter. 

1. What are the three properties of a vector, other than its contents?
  + length
  + type (?)
  + attributes (?)

1. What are the four common types of atomic vectors? What are the two
   rare types?
  + common: logical, character, numeric (double?) and integer
  + rare: complex and ???

1. What are attributes? How do you get them and set them?
  + attributes are "extra information" about an object, other than its contents
  + set via `attributes`

1. How is a list different from an atomic vector? How is a matrix different
   from a data frame?
  + a list can have elements of different types, a vector must have all of the same type
  + matrices must be all of the same type

1. Can you have a list that is a matrix? Can a data frame have a column
   that is a matrix?
  + no, though a list element could be a matrix (?)
  + yes, I think.
  
## Vectors

vectors are basic to R data types.  There are two kinds:

```{r}
x <- list("A")
typeof(x)
m <- lm(disp ~ cyl, data = mtcars)
typeof(m) 
length(m)
attributes(x)
attributes(m)

## this should be FALSE because of the class attribute:
is.vector(m)
## and this is TRUE:
is.vector(x)
is.vector(list(a = "A"))

## lists are vectors, but they're not atomic:
is.atomic(x)
```

## coercion

the heirarchy of "flexibility":

 logical < integer < double < character
 
 Makes sense: increases as the number of possible values
 
 2 < any integer < any number < anything you can type
 
 ## lists

I thought this example was interesting:

```{r}
x <- list(list(1, 2), c(3, 4))
y <- c(list(1, 2), c(3, 4))
str(x)
str(y)
```

`x` has the structure that I would expect -- ie it looks like what went in: one list, and one numeric vector.

`y` is a little counterintuitive: it just stuck everything in a single vector, which might actually come in handy:

```{r}
unlist(y)
data.frame(y)

data.frame(c(a = 1, b = 3))
data.frame(list(a = 1, b = 3))

```

If you coerce a name list of equal-length arguments, you get a pretty dataframe.  If you do so to a named vector, you get a single column with a coerced name. 

### Exercises

1. What are the six types of atomic vector? How does a list differ from an
   atomic vector?
  + logical, integer, double, character, (complex, raw)
  + lists can have elements of various types, various lengths, and are "recursive"

1. What makes `is.vector()` and `is.numeric()` fundamentally different to
   `is.list()` and `is.character()`?
  + `is.vector` returs `TRUE` for **any vector of any type** that doesn't have attributes besides names. either atomic or list.
  + `is.numeric`, similarly, is `TRUE` for either integer or double vectors, but not for lists.
```{r}
is.numeric(list(1,2,3))
is.numeric(c(1,2,3))
is.numeric(c(1L,2L,3L))
```
  + `is.list` is TRUE for only one of the two types of vectors, while `is.vector` is true for both
  + `is.character` is true for only one **type** of vector, while `is.numeric` is true for **2**
  
1. Test your knowledge of vector coercion rules by predicting the output of
   the following uses of `c()`:

    ```{r, eval=FALSE}
    c(1, FALSE)  # should be 1 0
    c("a", 1) # "a" "1"
    c(list(1), "a") # "1" "a"
    ## WRONG, it is a list, not a vector! eqiv to list(1,"a")
    c(TRUE, 1L) #1 1
    ```

1. Why do you need to use `unlist()` to convert a list to an
    atomic vector? Why doesn't `as.vector()` work?
  + because it is already a vector
  
```{r}
list("A")
as.vector(list("A"))
unlist(list("A"))
```

1. Why is `1 == "1"` true? Why is `-1 < FALSE` true? Why is `"one" < 2` false?
  + from the help:

>If the two arguments are atomic vectors of different types, one is coerced to the type of the other, the (decreasing) order of precedence being character, complex, numeric, integer, logical and raw.

1. Why is the default missing value, `NA`, a logical vector? What's special
   about logical vectors? (Hint: think about `c(FALSE, NA_character_)`.)
  + as this example illustrates, using `c` with a character NA coerces everyting to character. Logical is a the "bottom" of the heirarchy, so it can't drag anything else up to its level.
  
## exercises

1. What does `dim()` return when applied to a vector?
```{r}
dim(c(1,2)) # NULL?
```

1. If `is.matrix(x)` is `TRUE`, what will `is.array(x)` return?
```{r}
is.matrix(matrix(0))
is.array(matrix(0)) ## also true
```
1. How would you describe the following three objects? What makes them
    different to `1:5`?

    ```{r}
    x1 <- array(1:5, c(1, 1, 5))
    x2 <- array(1:5, c(1, 5, 1))
    x3 <- array(1:5, c(5, 1, 1))
    ```
    These are all 3D arrays with `length(dim()) = 3` . 
    
    
### Exercises

1. What attributes does a data frame possess?

```{r}
   foo <- data.frame(a = "A")
   foo
   attributes(foo)
```
names, row.names, and class


1. What does `as.matrix()` do when applied to a data frame with
    columns of different types?
  probably coerces them
  

1. Can you have a data frame with 0 rows? What about 0 columns?

What is "mode"?

```{r}
x <- 1:3
typeof(x)
mode(x)
```

