---
title: "Debugging"
author: "Melissa"
date: "August 6, 2014"
output: html_document
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE)
```


1. Compare the following two implementations of message2error(). What is the main advantage of withCallingHandlers() in this scenario? (Hint: look carefully at the traceback.)

```{r}
message2error <- function(code) {
  withCallingHandlers(code, message = function(e) stop(e))
}

message2error <- function(code) {
  tryCatch(code, message = function(e) stop(e))
}
```



2. The goal of the col_means() function defined below is to compute the means of all numeric columns in a data frame.

```{r}
col_means <- function(df) {
  numeric <- sapply(df, is.numeric)
  numeric_cols <- df[, numeric]

  data.frame(lapply(numeric_cols, mean))
}
```


However, the function is not robust to unusual inputs. Look at the following results, decide which ones are incorrect, and modify col_means() to be more robust. (Hint: there are two function calls in col_means() that are particularly prone to problems.)

```{r}
col_means(mtcars) #Good
colMeans(mtcars)

col_means(mtcars[, 0]) #Bad #fixed
colMeans(mtcars[, 0])

col_means(mtcars[0, ]) #Good ***
colMeans(mtcars[0, ])

col_means(mtcars[, "mpg", drop = F]) #Bad #fixed
colMeans(mtcars[, "mpg", drop = F])

col_means(1:10) #Good
colMeans(1:10)

col_means(as.matrix(mtcars)) #Bad 
colMeans(as.matrix(mtcars))

col_means(as.list(mtcars)) #Good
colMeans(as.list(mtcars)) 


mtcars2 <- mtcars
mtcars2[-1] <- lapply(mtcars2[-1], as.character)
col_means(mtcars2) #Bad
colMeans(mtcars2)

```


```{r}
#Not working for all :(
col_means <- function(df) {
  if(!dim(df)==2){stop('df must have at least two dimensions')}
numeric <- apply(df, 2, is.numeric)
  numeric_cols <- df[, numeric, drop=FALSE]
  apply(numeric_cols,2 , mean)
}

#Again not working for all :'(
col_means2 <- function(df) {
  numeric <- vapply(df, is.numeric, FUN.VALUE=1)
  numeric_cols <- df[, numeric, drop=FALSE]

  data.frame(lapply(numeric_cols, mean))
}
```



3. The following function “lags” a vector, returning a version of x that is n values behind the original. Improve the function so that it (1) returns a useful error message if n is not a vector, and (2) has reasonable behaviour when n is 0 or longer than x.

```{r}
lag <- function(x, n = 1L) {
  
  if(!is.numeric(n)) {stop('n is not a numeric')}
  if(n==0) {warning('n is 0, so it didnt lag')}
  if(n>length(x)){stop('n is larger than x!')}
  xlen <- length(x)
  c(rep(NA, n), x[seq_len(xlen - n)])
}

lag(1:10)
lag(1:10, 'a')
lag(1:10, 20)
lag(1:10, 0)
```








