---
title: "Functionals"
author: "Melissa"
date: "September 2, 2014"
output: html_document
---


```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE)
```

rm(list=ls())

**lapply**

```{r}
trims <- c(0, 0.1, 0.2, 0.5)
x <- rcauchy(1000)
unlist(lapply(trims, function(trim) mean(x, trim = trim)))
```

1. Why are the following two invocations of lapply() equivalent?

trims <- c(0, 0.1, 0.2, 0.5)
x <- rcauchy(100)

lapply(trims, function(trim) mean(x, trim = trim))
lapply(trims, mean, x = x)


2. The function below scales a vector so it falls in the range [0, 1]. How would you apply it to every column of a data frame? How would you apply it to every numeric column in a data frame?

scale01 <- function(x) {
  rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
}

```{r}
lapply(mtcars, FUN=scale01)
```
```{r}
mydf<-data.frame(letters,1:26)

scale01numeric <- function(x) {
  if(is.numeric(x)){
    rng <- range(x, na.rm = TRUE)
  (x - rng[1]) / (rng[2] - rng[1])
  }
}

lapply(mydf, FUN=scale01numeric)
```



3. Use both for loops and lapply() to fit linear models to the mtcars using the formulas stored in this list:

formulas <- list(
  mpg ~ disp,
  mpg ~ I(1 / disp),
  mpg ~ disp + wt,
  mpg ~ I(1 / disp) + wt
)

```{r}
res<-list()

for(i in 1:length(formulas)){
  res[[i]]<-lm(formulas[[i]], data=mtcars)
  }
res

```

```{r}
lapply(formulas, FUN=function(x) lm(formula=x, data=mtcars))
```



4. Fit the model mpg ~ disp to each of the bootstrap replicates of mtcars in the list below by using a for loop and lapply(). Can you do it without an anonymous function?

bootstraps <- lapply(1:10, function(i) {
  rows <- sample(1:nrow(mtcars), rep = TRUE)
  mtcars[rows, ]
})

```{r}

res2<-list()
for(i in 1:length(bootstraps)){
  res2[i]<-lm(mpg~disp, data=bootstraps[[i]])
}
#not sure why its not working?? ---- double square brakets!
res2
```

```{r}
lapply(bootstraps, FUN=function(x) lm(mpg~disp, data=x))

```



5. For each model in the previous two exercises, extract R2 using the function below.

rsq <- function(mod) summary(mod)$r.squared

```{r}
lapply(formulas, FUN=function(x) rsq(lm(formula=x, data=mtcars)))
```


```{r}
lapply(bootstraps, FUN=function(x) rsq(lm(mpg~disp, data=x)))

```


**friends of lapply**


1. Use vapply() to:

  a. Compute the standard deviation of every column in a numeric data frame.
  
```{r}
vapply(mtcars, FUN=sd, numeric(1))
```

  b. Compute the standard deviation of every numeric column in a mixed data frame. (Hint: you’ll need to use vapply() twice.)
  
```{r}
mydf2<-data.frame(rep('a',26),1:26,2:27)

mydf.numeric<-vapply(mydf2, FUN=is.numeric, FUN.VALUE=1)

vapply(mydf.numeric,FUN=function(x) sd(mydf2[,x]),FUN.VALUE=1)

```


2. Why is using sapply() to get the class() of each element in a data frame dangerous?

Because sapply will work even if class has more than one value. 

3. The following code simulates the performance of a t-test for non-normal data. Use sapply() and an anonymous function to extract the p-value from every trial.

trials <- replicate(
  100, 
  t.test(rpois(10, 10), rpois(7, 10)),
  simplify = FALSE
)
Extra challenge: get rid of the anonymous function by using [[ directly.

```{r}
sapply(trials, function(x) x$ p.value)

```


4. What does replicate() do? What sort of for loop does it eliminate? Why do its arguments differ from lapply() and friends?

it evaluates an expression n number of times. It differs because it evaluates expressions and not just functions. 

```{r}
replicate(3,'dont stop')
```


5. Implement a version of lapply() that supplies FUN with both the name and the value of each component.

6. Implement a combination of Map() and vapply() to create an lapply() variant that iterates in parallel over all of its inputs and stores its outputs in a vector (or a matrix). What arguments should the function take?

7. Implement mcsapply(), a multicore version of sapply(). Can you implement mcvapply(), a parallel version of vapply()? Why or why not?




**apply and sweep**

1. How does apply() arrange the output? Read the documentation and perform some experiments.

It reduces the number of dimensions by one (by the margin given) , giving either a vector or a matrix

```{r}
a<-matrix(1:20, nrow=5)
apply(a, MARGIN=2, sum)
```


2. There’s no equivalent to split() + vapply(). Should there be? When would it be useful? Implement one yourself.

```{r}

pulse <- round(rnorm(22, 70, 10 / 3)) + rep(c(0, 5), c(10, 12))
group <- rep(c("A", "B"), c(10, 12))
tapply(pulse, group, mean)

tapply3 <- function(x, group, f, ..., output) {
  pieces <- split(x, group)
  vapply(pieces, f, FUN.VALUE=output)
}

tapply3(pulse, group, mean, output=1)

```

3. Implement a pure R version of split(). (Hint: use unique() and subsetting.) Can you do it without a for loop?

```{r}
pulse <- round(rnorm(22, 70, 10 / 3)) + rep(c(0, 5), c(10, 12))
group <- rep(c("A", "B"), c(10, 12))
pieces <- split(pulse, group)
```

```{r}
unique(group)
split2<-function(vector, group){
  types<-unique(group)
  numgroup<-lapply(types,FUN= function(x) sum(group==x))
  lapply(numgroup, function(lev) )
  subset(pulse,1:10)
}

  
```


4. What other types of input and output are missing? Brainstorm before you look up some answers in the plyr paper.

output: data frame



**Manipulating lists**

1. Why isn’t is.na() a predicate function? What base R function is closest to being a predicate version of is.na()?

v<-c(NA)
is.na(v)

2. Use Filter() and vapply() to create a function that applies a summary statistic to every numeric column in a data frame.

```{r}

numsum<-function(df1){
  dob<-Filter(is.numeric, df1)
  vapply(dob, FUN=summary, FUN.VALUE=double(6))
}

numsum(mtcars)

```


3. What’s the relationship between which() and Position()? What’s the relationship between where() and Filter()?

which is the general form of position. where is the general form of filter? where will show you the true and false, filter only shows the true. 

4. Implement Any(), a function that takes a list and a predicate function, and returns TRUE if the predicate function returns TRUE for any of the inputs. Implement All() similarly.


```{r}

lt<-list(1:3,5:6,c('a','b','c'))

any(Filter(is.numeric,lt))
```



5. Implement the span() function from Haskell: given a list x and a predicate function f, span returns the location of the longest sequential run of elements where the predicate is true. (Hint: you might find rle() helpful.)


```{r}

span<-function(x,f,condition){
  f(condition, x)
  rle(f)
}
Filter(is.numeric, lt)

```



**Mathematical Functionals**

1. Implement arg_max(). It should take a function and a vector of inputs, and return the elements of the input where the function returns the highest value. For example, arg_max(-10:5, function(x) x ^ 2) should return -10. arg_max(-5:5, function(x) x ^ 2) should return c(-5, 5). Also implement the matching arg_min() function.

```{r}

arg_max<-function(inter, fun){
  ap<-optimise(fun, inter, maximum=TRUE)$maximum
  v<-round(ap)
  inter[which(v==inter | abs(v)==inter)]
}


arg_max(-10:5, function(x) x^2)
arg_max(-5:5, function(x) x ^ 2)

arg_min<-function(inter, fun){
  ap<-optimise(fun, inter)$minimum
  v<-round(ap)
  inter[which(v==inter | abs(v)==inter)]
}
arg_min(-10:5, function(x) x^2)

```


2. Challenge: read about the fixed point algorithm. Complete the exercises using R.

```{r}
fixed.point<-function(fun, guess0=100){
  tolerance<-0.001
  guess<-rnorm(20,guess0,guess0) 
  fofguess<-sapply(guess,FUN=fun)
  dif1<-fofguess-guess
  mdif<-min(abs(dif1))
  guess1<-guess[which(dif1==mdif | dif1==-mdif)]
  if(mdif< tolerance){return(guess1)}
  else{
    fixed.point(fun, guess0=guess1)
  }
}

fixed.point(fun=cos)
fixed.point(fun= function(x) 1+ (1/x))
```




3. How does paste() fit into this structure? What is the scalar binary function that underlies paste()? What are the sep and collapse arguments to paste() equivalent to? Are there any paste variants that don’t have existing R implementations?
paste-> 




