---
title: "Functions"
author: "Alathea"
date: '2014-07-16'
output: 
  html_document:
    keep_md: yes
    toc: yes
---

***

## Exercises

***

## Reading Notes

### Quiz

#### What are the three components of a function?

body, formals, environment

#### What does the following code return?

```{r quiz1, eval = FALSE}
y <- 10
f1 <- function(x) {
  function() {
    x + 10
  }
}
f1(1)()
```

11

#### How would you more typically write this code?
**`` `+`(1, `*`(2, 3))``**

I think `1 + (2 * 3)`

#### How could you make this call easier to read?
**`mean(, TRUE, x = c(1:10, NA))`**



#### Does the following function throw an error when called?  Why/why not?
```{r quiz2, eval = FALSE}
f2 <- function(a, b) {
  a * 10
}
f2(10, stop("This is an error!"))
```



#### What is an infix function? How do you write it? What’s a replacement function? How do you write it?



#### What function do you use to ensure that a cleanup action occurs regardless of how a function terminates?


### Function components

body, formals, environment

functions can have a custom print method (as in modeling functions)

primitive functions do not have the three components; different methods are used to match arguments:
* `switch` = chooses an argument based on a given expression
* `call`

***

## Discussion Notes

***