---
title: "Functions"
author: "Melissa"
date: "16 July, 2014"
output: html_document
---
```{r, echo=FALSE}
rm(list=ls())
```


#Quiz#

1. What are the three components of a function?
 NI

2. What does the following code return? 11

```{r}
y <- 10
f1 <- function(x){
  function(){
    x + 10
  }
}

f1(1)()
```

3. How would you more typically write this code?

`+`(1, `*`(2, 3))

```{r}
1 + 2*3

```

4. How could you make this call easier to read?

mean(, TRUE, x = c(1:10, NA))

```{r}
mean(1:10, na.rm=TRUE)
```


5. Does the following function throw an error when called? Why/why not?
  No because the items needed inside the function (a) were provided. Since b is not needed, it is not used

```{r}
f2 <- function(a, b) {
  a * 10
}
f2(10, stop("This is an error!"))
#f2(stop('this is an error'),10) -> it shows an error
```


6. What is an infix function? How do you write it? What’s a replacement function? How do you write it?

NI

7. What function do you use to ensure that a cleanup action occurs regardless of how a function terminates?

NI

#Excercises#

1. What function allows you to tell if an object is a function? What function allows you to tell if a function is a primitive function?

```{r}
is.function(sum)
is.primitive(sum)

is.function(f1)
is.primitive(f1)

```

2. This code makes a list of all functions in the base package.

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
```

Use it to answer the following questions:

Which base function has the most arguments?

```{r}

arg_num<-sapply(funs,FUN=function(x){length(formals(x))})
max(arg_num)
which(arg_num==21)

```


How many base functions have no arguments? What’s special about those functions.

```{r}
length(which(arg_num==0))

```


How could you adapt the code to find all primitive functions?

```{r}
primitive_test<-lapply(funs,FUN=function(x){is.primitive(x)})
which(primitive_test==TRUE)

```


3. What are the three important components of a function?

The environment, the formals and the body. 

4. When does printing a function not show what environment it was created in?

It means that the function was created in the global environment. 



1. What does the following code return? Why? What does each of the three c’s mean?
 -c1 is a object (10)
 -c2 is a function (concatenate)
 -c3 is a name (c)
 
```{r}
c <- 10
c(c = c)
```

2. What are the four principles that govern how R looks for values?
 - If a name isn’t defined inside a function, R will look one level up
 - Finding functions works exactly the same way as finding variables
 - Every time a function is called, a new environment is created to host execution
 - R looks for values when the function is run, not when its created
 

3. What does the following function return? Make a prediction before running the code yourself.
 202
```{r}
f <- function(x) {
  f <- function(x) {
    f <- function(x) {
      x ^ 2
    }
    f(x) + 1
  }
  f(x) * 2
}
f(10)
```

rm(list=ls())
data.frame(x=1:10,y=x*2)



1. Clarify the following list of odd function calls:

```{r}
x <- sample(replace = TRUE, 20, x = c(1:10, NA))
x<- sample(1:10, 20, replace=TRUE)
y <- runif(min = 0, max = 1, 20)
y<- runif(20)
cor(m = "k", y = y, u = "p", x = x)
cor(x, y, method='kendall', use='pairwise.complete.obs')
```

2. What does this function return? Why? Which principle does it illustrate?
3, lazy evaluation for y=0 and dynamic lookup in the sense that values are only evaluated when used

```{r}
f1 <- function(x = {y <- 1; 2}, y = 0) {
  x + y
}
f1()
```

3. What does this function return? Why? Which principle does it illustrate?
100. Default arguments are evaluated inside the function. This means that if the expression depends on the current environment the results will differ depending on whether you use the default value or explicitly provide one. Dynamic look up, it looks it up in the environment of the functions body and then it would look in the global environment. 

```{r}
f2 <- function(x = z) {
  z <- 100
  x
}
f2()
```




1. Create a list of all the replacement functions found in the base package. Which ones are primitive functions?

```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
repfun<-funs[agrep('<-',funs)]

??agrep for '<-'

```

2. What are valid names for user created infix functions?

%something%

3.Create an infix xor() operator.
rm(list=ls())

```{r}
'%or%'<-function(x,y){
  if((x== TRUE & y== FALSE) | (x==FALSE & y== TRUE)) {TRUE}  
  else {FALSE}
  }

xor(0,1)
0 %or% 1

'%orDiane%'<-function (a,b) a!=b


```

4. Create infix versions of the set functions intersect(), union(), and setdiff().


```{r}
rm(list=ls())

y<-1:10
x<-3:15


'%intersect%'<-function(x,y){
  x[x %in% y]
}

x %intersect% y
intersect(x,y)

'%union%'<-function(x,y){
  unique(c(x,y))
}

y %union% x
union(y,x)

'%setdiff%'<-function(x,y){
  x[!(x %in% y)]
}

setdiff(y,x)
y %setdiff% x
```


5. Create a replacement function that modifies a random location in a vector.

data.table!! use it for very big data. Good for copying things faster. Manages very big data!

```{r}
`ranreplacement<-` <- function(x, value){
  x[sample(length(x),1)]<-value
  x
}

x<-1:15
ranreplacement(x)<-25
x
```





1. How does the chdir parameter of source() compare to in_dir()? Why might you prefer one approach to the other?

chdir will actually change the working directory temporarily, whereas the in_dir will not. If you need to do other things like save in that other wd, then chdir might be more useful than indir()

2. What function undoes the action of library()? How do you save and restore the values of options() and par()?

detach-> library
foo<-par() - par(foo)


3. Write a function that opens a graphics device, runs the supplied code, and closes the graphics device (always, regardless of whether or not the plotting code worked).

on.exit(dev.off())
force(code)

4. We can use on.exit() to implement a simple version of capture.output().

rm(list=ls())

capture.output2 <- function(code) {
  temp <- tempfile()
  on.exit(file.remove(temp), add = TRUE)

  sink(temp)
  on.exit(sink(), add = TRUE)

  force(code)
  readLines(temp)
}
capture.output2(cat("a", "b", "c", sep = "\n"))
#> [1] "a" "b" "c"

capture.output

capture.output(cat("a", "b", "c", sep = "\n"))
You might want to compare this function to the real capture.output() and think about the simplifications I’ve made. Is the code easier to understand or harder? Have I removed important functionality?

5. Compare capture.output() to capture.output2(). How do the functions differ? What features have I removed to make the key ideas easier to see? How have I rewritten the key ideas to be easier to understand?

help




**** use on.exit()

