---
title: "functions-Andrew.Rmd"
author: "Andrew MacDonald"
date: '2014-07-17'
output:
  html_document:
    keep_md: TRUE
---

```{r message=FALSE}
library(dplyr)
library(magrittr)
library(pryr)
```


1.  What function allows you to tell if an object is a function? What function
    allows you to tell if a function is a primitive function?


```{r}
is.function(ls)
is.primitive(ls)
is.primitive(c)
is.primitive(mean)
```


1.  This code makes a list of all functions in the base package. 
    
    ```{r}
    objs <- mget(ls("package:base"), inherits = TRUE)
    funs <- Filter(is.function, objs)
    ```

    Use it to answer the following questions:

    1. Which base function has the most arguments?

```{r}
which.max(sapply(funs,function(x) length(formals(x))))
```

    
    1. How many base functions have no arguments? What's special about those
       functions.
       
```{r}
funs[which(sapply(funs,function(x) length(formals(x)))==0)] %>% length

funs[which(sapply(funs,function(x) length(formals(x)))==0)] %>% sapply(is.primitive) %>% sum

funs[which(sapply(funs,function(x) length(formals(x)))==0)] %>% lapply(formals)

c <- 10
c(c = c)

```       
       
    1. How could you adapt the code to find all primitive functions?
```{r}
Filter(is.primitive, objs)
```


1. What are the three important components of a function?

*formals, body, environment*

1. When does printing a function not show what environment it was created in?

*When it is made in the base environment.*

***

1. What does the following code return? Why? What does each of the three `c`'s mean?

    ```{r, eval = FALSE}
    c <- 10
    c(c = c)
    ```

*1st is a function, 2nd a formal, 3rd a variable*

2. What are the four principles that govern how R looks for values?

*name masking* -- use the most recent assignment to a name, in the "closest" environment
*functions vs. variables* -- same rules for functions and variables. additionally, R knows when something is a function from context.
*a fresh start* -- functions spawn new, clean environments to execute.  they don't remember previous runs.
*dynamic lookup* -- look for values **now**, in the present environments. **BE SO CAREFUL**, since if you don't say what your function uses, you are assuming that those other variables don't change.

3. What does the following function return? Make a prediction before 
   running the code yourself.

    ```{r, eval = FALSE}
    f <- function(x) {
      f <- function(x) {
        f <- function(x) {
          x ^ 2
        }
        f(x) + 1
      }
      f(x) * 2
    }
    f(10)
    ```

PREDICTION: 202
(weird -- distorting order of operations with closures! perverse)

1.  Clarify the following list of odd function calls:

    ```{r, eval = FALSE}
#x <- sample(replace = TRUE, 20, x = c(1:10, NA))
x  <- 1:10 %>%
  c(NA) %>%
  sample(20, replace = TRUE)

#y <- runif(min = 0, max = 1, 20)
y <- runif(20,0,1)

#cor(m = "k", y = y, u = "p", x = x)
cor(x, y, method = "kendall", use = "pairwise")

    ```


1.  What does this function return? Why? Which principle does it illustrate?
  
    ```{r, eval = FALSE}    
    f1 <- function(x = {y <- 1; 2}, y = 0) {
      x + y
    }
    f1()
    ```

*name masking: x is set to 2, and y is set to 1, once the function body is run.  `y = 0` is not kept.*
  
    ```{r, eval = FALSE}    
    f1 <- function(x = {y <- 1; 2}, y = 0) {
      print(y)
      print(x + y)
      print(y)
    }
    f1()
    ```

1.  What does this function return? Why? Which principle does it illustrate?

    ```{r, eval = FALSE}
    f2 <- function(x = z) {
     
      z <- NULL
      x
    }
    f2()
    ```

*er, also lazy evaluation? `x` is defined in terms of an object that exists only in the function body -- which is OK, since R doesn't look for `x` until you ask for it* 
* Dynamic lookup* 


## adders the closures:

```{r}
add <- function(x) {
  function(y) x + y
}
adders <- lapply(1:10, add)
adders[[1]](10)
#> [1] 20
adders[[10]](10)
#> [1] 20
```

OK, so in the closures created by `add`, x is a "promise".  We can see that all these "promises" have the same environment:

```{r}
addpromise <- function(x) {
  print(promise_info(x))
  function(y) x + y
}

adders <- lapply(1:10, addpromise)
adders[[1]](10)
sapply(adders, environment) %>% sapply(parent.env)
```

## modify in place

```{r}
foo <- 1:10
class(foo)
address(foo)
foo[2] <- 5
address(foo)
class(foo)

foo <- 1:10
class(foo)
address(foo)
foo[2] <- 5L
class(foo)
address(foo)

```

only works if you keep classes the same!

## Exercises

1. Create a list of all the replacement functions found in the base package. 
   Which ones are primitive functions?
   
```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
replacers <- funs %>% 
  names %>% 
  grepl("<-",.) %>% 
  funs[.] 

replacers[sapply(replacers, is.primitive)]
```

2. What are valid names for user created infix functions?

`%x%`, where ` x` is anything you can type (except `%`)

3. Create an infix `xor()` operator.

```{r}
`%xor%` <- function(a, b) {
  if(sum(a,b) == 1) TRUE else FALSE
  }

TRUE %xor% FALSE
TRUE %xor% TRUE


`%xorDiane%` <- function(a,b)  a != b

TRUE %xorDiane% FALSE

5789 %xorDiane% "Moose"

```

4. Create infix versions of the set functions `intersect()`, `union()`, and 
   `setdiff()`.

```{r}
`%inter%` <- function(a,b) intersect(a,b)

letters[1:5] %inter% letters[3:7]
```

5. Create a replacement function that modifies a random location in a vector.

```{r}
`randomod<-` <- function(x, value) {
  position <- sample(seq_along(x),size = 1)
  x[position] <- value
  x
}

bar <- 1:5 
randomod(bar) <- 500
bar
```

