---
title: "Functions"
author: "Susannah"
output: 
  html_document:
    keep_md: yes
---

## Quiz 
1. The three elements of a function are the name of the function, the arguments of the function, and the guts of the function. *X* _Functions don't have to have names in `R`, but the environment matters. Remember that time Hadley's plyr masked a function in another package and you got really confused?_
2. 11
3. `1 + 2 * 3`
4. `mean(x = c(1:10, NA), na.rm = TRUE)`
5. No - the `stop` should be within the function, not the arguments to the function.
6. An infix function is a function like `+,-,/` that generally goes inbetween arguments. I don't know what a replacement function is. *X*
7. `on.exit()`

## Functional components 
1. `class` allows you to tell if an `R` object is a function. I tried to use `attributes()` to find out whether functions were primitive or not, but returned `NULL` for many functions - including functions I knew to be not primitive. Just typing the name of the function without parentheses will tell you whether the function is primitive or not. `environment(function)` returns `NULL` for all primitive functions.
2. 
```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
nullenv <- function(func) {
    is.null(environment(func))
}
funs <- Filter(is.function, objs)
nulls <- Filter(nullenv, funs)

arg <- lapply(funs, formals) # get formals
counts <- sapply(arg, length) # count formals
maxcount <- which.max(counts) #find most args
zeros <- which(counts == 0) #no args
```
  * `scan` has the most arguments - `r maxcount`.
  * infixes, math functions, casts, control flow, system calls, and debugging functions have no arguments. These are typically primitive functions (see output of `sapply(names(zeros),environment`). Since primitive functions handle arguments a bit differently than other functions, I guess it's best to avoid the problem altogether by not having arguments at all!
  * 
```{r}
  objs <- mget(ls("package:base"), inherits = TRUE)
  
  nullenv <- function(func) {
    # returns true if function environment is null
    is.null(environment(func))
  }
  
  funs <- Filter(is.function, objs)
  nulls <- Filter(nullenv, funs)
```
  
## Lexical scoping 

1. 
```{r} c <- 10
  c(c = c)
  ```
  This code returns a vector of length one containing the number 10. The vector has a names attribute of "c." The first `c` is the concatenate function. The second c is an attribute assignment and the third c is a variable, which was assigned to 10 in the previous line.
2. The four principles are 
  * name masking: names from the inner environment are used before those in the outer environment
  * functions vs. variables: R can distinguish between functions and variables, but it's confusing to _you_ if you have variables and functions with the same names
  * a fresh start: a new environment is created every time a function is run
  * dynamic lookup: R looks for values when the function is run, of course. This isn't a language where you have to compile things, so I don't think there's another way to do it...
3. 202. Evaluate inner to outer.
  
## EOiaFC 
> Everything that exists is an object.
> Everything that happens is a function call.
-- John Chambers

## Function Arguments 

Lazy evaluation means that things aren't evaluated unless you use them.
I don't understand what closures are, _e.g._ "This is important when creating closures with lapply or a loop"

This example is confusing. Let's work through it.
```{r}
add <- function(x) {
  function(y) x + y
}
adders <- lapply(1:10, add) 
adders[[1]](10)
```

lapply step creates a list of functions and _environments_ where x is assigned according to the inputs to `lapply`. 

```{r}
ls(environment(adders[[1]]))
```
You'd think this would lead to x being 1 in the first environment, 2 in the second environment and so on. However, because of lazy evaluation, `x` is `10` in all environments. 

```{r}
for (i in 1:10) {
  print(get("x", pos=environment(adders[[i]])))
}
```

I think this has to do with promises and the way `lapply` works. `lapply` is basically a loop in C that in this case is like

`for (i in 1:10) {
  adders[[i]] <- add(i)
}`

This loop executes in its own environment where a promise gets made for x at every iteration. At no point have we been asked to evaluate `x`, so no value has been assigned to it - just the [expression `x` and a pointer to the environment the function was called from](http://cran.r-project.org/doc/manuals/r-release/R-lang.html#Promise-objects). At the end of the `lapply` loop when the promises are due, `x` is `10`, so `x` becomes `10` in all of the `adders` environments. 

If we `force(x)`, we cause the promises to come due _before_ the loop ends, so that x gets assigned to the `i` in the `for` loop above instead of just the last number in the sequence. 

```{r}
add <- function(x) {
  force(x)
  function(y) x + y
}
adders <- lapply(1:10, add) 
adders[[1]](10)
for (i in 1:10) {
  print(get("x", pos=environment(adders[[i]])))
}
```

1. 
```{r eval=FALSE}
x <- sample(c(1:10, NA), 20, replace = TRUE)
y <- runif(20, min = 0, max = 1)
cor(x, y, method = "kendall", use = "pairwise.complete.obs")
```
2. The function returns 3. It illustrates default arguments and lexical scoping (name masking).
3. This function returns 100. It illustrates promises/lazy evaluation. 

# Special Calls
1.
```{r}
objs <- mget(ls("package:base"), inherits = TRUE)
funs <- Filter(is.function, objs)
repfun <- funs[grep(".*<-", funs)] # List of all replacement functions
nulls <- Filter(nullenv, repfun)
```
Primitive replacement functions are `r names(nulls)`
2. Valid names for infix functions begin and end with `%`.
3. 
```{r}
`%|||%` <- function(x, y) {
  ( x | y) & !(x & y) # defn on wikipedia
}
x <- c(1, 2, 3)
y <- c(1, 2, 3)
(x > y) %|||%  (y < x)
(x == y) %|||% (x > y)
```
4.
```{r}
`%uU%` <- function(x, y) {
    # find intersection of two vectors
    matches <- match(x, y, nomatch = 0)
    unique(y[matches])
}
x <- seq(2, 50, 3)
y <- seq(1, 50, 2)

all(x %uU% y == intersect(x, y))

`%U%` <- function(x, y) {
    # find the union of two vectors
    unique(c(x,y))
}

all(x %U% y == union(x,y))


`%!!%` <- function(x, y) {
    # find differences between two vectors
    matches <- match(x, y, nomatch = 0)
    unique(x[matches==0])
}

all(x %!!% y == setdiff(x, y))
```
5.
```{r}
`modify_rand<-` <- function(x, value) {
  position <- floor(runif(1, max = length(x)))
  x[position] <- value
  x
}
modify_rand(x) <- 10
x
```
# Return Values

I don't understand what `on.exit()` does when executed without an expression.

1. If `chdir` is set to `TRUE`, the `R` working directory is temporarily changed to the directory containing the source file. `in_dir()` allows you to set the directory to a new location, not necessarily the location of the source code, providing more flexibility.
2. `detach()` undoes the action of `library()`. To change the values of `options()` and `par()` do something like `options(name = value)`.  Before you go changing individual name-value pairs, assign the entire list of options or parameters to some object so you'll be able to reset it, like so:
```{r}
newop <- options() # assign options to a new list
options(verbose=TRUE) # talk to me R! # save a new value of an option
options(newop) # restore all options
options(newop[which(names(newop) == "verbose")]) #restore just the verbose option
```
3. My solution doesn't work
```{r}
make_plot <- function(file, plot_code, width=500, height=500) {
  png(file, width=width, height=height)
  plot_code
  graphics.off()
}

simpleplot <- plot(x = c(1:10), y = rnorm(10))
#borked <- plot(x = c(1:9), y = rnorm(10))

#make_plot("~/Documents/classes/UBCadv-r/03_Functions/Functions_Susannah_plot.png", borked)
make_plot("~/Documents/classes/UBCadv-r/03_Functions/Functions_Susannah_plot.png", simpleplot)
```
4. Code is easier to understand. Output can't be directed to a specific file with `capture.output2`. Lose access to errors.
5.
```{r}
capture.output <- function(..., file=NULL, append=FALSE)
{
    args <- substitute(list(...))[-1L]

    rval <- NULL; closeit <- TRUE
    if (is.null(file))
        file <- textConnection("rval", "w", local = TRUE) #if no file provided make a textConnection
    else if (is.character(file))
        file <- file(file, if(append) "a" else "w")
    else if (inherits(file, "connection")) {
if (!isOpen(file)) open(file, if(append) "a" else "w")
else closeit <- FALSE
    } else
        stop("'file' must be NULL, a character string or a connection")

    sink(file) # divert output to the file
    ## for error recovery: all output will be lost if file=NULL
    on.exit({sink(); if(closeit) close(file)})

    pf <- parent.frame() 
    evalVis <- function(expr)
        withVisible(eval(expr, pf))

    for(i in seq_along(args)) {
        expr <- args[[i]]
        tmp <- switch(mode(expr),
                      "expression" = lapply(expr, evalVis),
                      "call" =, "name" = list(evalVis(expr)),
                       stop("bad argument")) 
        for(item in tmp)
            if (item$visible) print(item$value)
    }
    ## we need to close the text connection before returning 'rval'
    on.exit()
    sink() # I don't get why there are multiple sinks and on.exits
    if(closeit) close(file)
    if(is.null(rval)) invisible(NULL) else rval #how could rval be anything _but_ NULL in this function?
}
```
I don't have a good understanding of what either function is doing.

