---
title: "Subsetting"
author: "Melissa"
date: "10 July, 2014"
output: html_document
---


```{r, echo=FALSE}
rm(list=ls())
```

**1.Fix each of the following common data frame subsetting errors:**

```{r}
     
mtcars[mtcars$cyl==4,]
                
mtcars[-(1:4),]
     
mtcars[mtcars$cyl <= 5,]

mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ] 
mtcars[mtcars$cyl %in% c(4,6),]
```
%in% maps the second thing to the first things, like match. 

**2. Why does x <- 1:5; x[NA] yield five missing values? Hint: why is it different from x[NA_real_]?**

Logical vectors are recycled. So NA being a logical vector gets recycled five times. When NA is real, it doesnt get recycled. And a missing value on the index returns a missing value on the output.
```{r}
  x<-1:5 
  x[NA]
  x[NA_real_]
```
  
**3. What does upper.tri() return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?**

```{r}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```
It returns a logical matrix of the the upper triangle of the matrix (ignoring the diagonal). Subsetting with it returns an atomic vector of the TRUE values in a column-major order.


**4.Why does mtcars[1:20] return a error? How does it differ from the similar mtcars[1:20, ]?**

```{r}
mtcars[1:20,]
nrow(mtcars)
ncol(mtcars)
```
Too many columns, mtcars[1:20] calls for the columns, so in this case there are too many columns. mtcars[1:20,] calls the first 20 rows in all the columns 

**5.Implement your own function that extracts the diagonal entries from a matrix (it should behave like diag(x) where x is a matrix).**

```{r}
x<-matrix(1:9,ncol=3)
y<-matrix(c(1:ncol(x),1:nrow(x)),nrow=ncol(x))
x[y]
diag(x)
```
Sorry not a function :(

**6.What does df[is.na(df)] <- 0 do? How does it work?**

It replaces NA in the data frame by 0. is.na(df) creates a logical df with TRUE and FALSE, and then df[] selects the values that returned TRUE



**7.Given a linear model, e.g. mod <- lm(mpg ~ wt, data = mtcars), extract the residual degrees of freedom. Extract the R squared from the model summary (summary(mod))**

```{r}
mod<-lm (mpg~wt, data=mtcars)
str(summary(mod),give.attr=FALSE)
summary(mod)$r.squared
summary(mod)$df.residual
```

8.
How would you randomly permute the columns of a data frame? (This is an important technique in random forests). Can you simultaneously permute the rows and columns in one step?




```{r}
df <- data.frame(x = rep(1:3, each = 2), y = 6:1, z = letters[1:6],w=8:13)

df

sample(ncol(df))
df[sample(nrow(df)),sample(ncol(df))]

```




9.
How would you select a random sample of m rows from a data frame? What if the sample had to be contiguous (i.e. with an initial row, a final row, and every row in between)?

```{r}
df <- data.frame(x = rep(1:3, each = 2), y = 6:1, z = letters[1:6],w=8:13)

sample(nrow(df),2)

```


10.
How could you put the columns in a data frame in alphaetical order?


```{r}
df <- data.frame(a = rep(1:3, each = 2), b = 6:1, c = letters[1:6],d=8:13)

df1<-df[,sample(ncol(df))]

df1[order(names(df1))]

```



