---
title: "Subsetting"
author: "Andrew MacDonald"
date: '2014-07-09'
<<<<<<< HEAD
output: md_document
---
```{r}
library(dplyr)
library(tidyr)
library(magrittr)
=======
output: 
  html_document:
    keep_md: yes
    theme: united
---
```{r message=FALSE}
library(dplyr)
library(tidyr)
library(magrittr)
library(ggplot2)
>>>>>>> upstream/master
```

## Exercises 1

1.  Fix each of the following common data frame subsetting errors:

    ```{r, eval = FALSE}
    #mtcars[mtcars$cyl = 4, ]
    mtcars[mtcars$cyl == 4, ]
    #mtcars[-1:4, ]
    mtcars[-(1:4), ]
   # mtcars[mtcars$cyl <= 5]
    mtcars[mtcars$cyl <= 5,]
#     mtcars[mtcars$cyl == 4 | 6, ]
    mtcars[mtcars$cyl == 4 |mtcars$cyl ==  6, ]
    ```

1.  Why does `x <- 1:5; x[NA]` yield five missing values? Hint: why is 
    it different from `x[NA_real_]`?
    
**Because `NA` is a logical vector, and therefore is recycled to the length of `x` (i.e. 5)**

```{r}
x <- 1:5; x[NA] ; x[NA_real_]
```

1.  What does `upper.tri()` return? How does subsetting a matrix with it 
    work? Do we need any additional subsetting rules to describe its behaviour?

    ```{r, eval = FALSE}
    x <- outer(1:5, 1:5, FUN = "*")
    x[upper.tri(x)]
    ```

**`upper.tri()` makes a logical matrix of class ``r upper.tri(x) %>% typeof``, which is used to subset x (same size and dimensions.  We don't need new subsetting rules because matrices are vectors!**

1.  Why does `mtcars[1:20]` return a error? How does it differ from the 
    similar `mtcars[1:20, ]`?
    
    **Because there are 11, not 20, columns in `mtcars` and this code is asking for the first 20 columns using vector notation. `mtcars[1:20,]` asks for the first 20 rows, which is no problem because there are 30.**

1.  Implement your own function that extracts the diagonal entries from a
    matrix (it should behave like `diag(x)` where `x` is a matrix).
    
```{r}
x  <-  outer(1:5, 1:5, FUN = "*")

andrew_diag <- function(mat){
  require(dplyr)
  require(magrittr)
  
  mat %>%
    ncol %>%
    seq_len %>%
    l(x -> cbind(x,x)) %>%
    mat[.]
  
  }

andrew_diag(x)

```


1.  What does `df[is.na(df)] <- 0` do? How does it work?

**uses vector notation to subset a dataframe, as if it were a matrix, and replaces all NA values with 0**

## Exercises 2

1.  Given a linear model, e.g. `mod <- lm(mpg ~ wt, data = mtcars)`, extract
    the residual degrees of freedom. Extract the R squared from the model
    summary (`summary(mod)`)

```{r}
mod <- lm(mpg ~ wt, data = mtcars)
summary(mod)[[c("fstatistic", "dendf")]]

<<<<<<< HEAD
## here is a little randomization test:
replicate(n = 20,{data.frame(x = runif(n = 20,10,20)) %>%
=======
data.frame(x = runif(n = 20,10,20)) %>%
  mutate(y = x * 2 + rnorm(20)) %>%
  qplot(x = x, y = y, data = .,geom = "point")

## here is a little randomization test:
replicate(n = 200,{data.frame(x = runif(n = 20,10,20)) %>%
>>>>>>> upstream/master
                    mutate(y = x * 2 + rnorm(20)) %>%
                    lm(y ~ x, data = .)
},simplify = FALSE
) %>% 
  lapply(summary) %>%
<<<<<<< HEAD
  sapply("[[",i = c("fstatistic", "value"))
=======
  sapply("[[",i = c("fstatistic", "value")) %>%
  qplot(geom = "density",xlab = "fstatistics")

>>>>>>> upstream/master

```

# Exercises 3

### Exercises

1.  How would you randomly permute the columns of a data frame? (This is an
    important technique in random forests). Can you simultaneously permute 
    the rows and columns in one step?

```{r}
df <- data.frame(a = 1:5,
                 b = 6:10,
                 c = 11:15)

df[sample(ncol(df))]
df[,sample(ncol(df))]
df[sample(nrow(df)),sample(ncol(df))]

```


1.  How would you select a random sample of `m` rows from a data frame? 
    What if the sample had to be contiguous (i.e. with an initial row, a 
    final row, and every row in between)?
    
```{r}
contin_samp <- function(df,m){
  df %>% 
    nrow %>% 
    subtract(m) %>% 
    sample(1) %>%
<<<<<<< HEAD
    add(seq_len(m)) %>% 
    df[.,]
=======
    add(0:m) %>% 
    df[.,,drop = FALSE]
>>>>>>> upstream/master
  }

data(mtcars)

contin_samp(mtcars,4)

<<<<<<< HEAD
testdata <- data.frame(a = 1:100, b = 2:101)
=======
testdata <- data.frame(a = 1:100)
>>>>>>> upstream/master

contin_samp(testdata,5)
```

    
1.  How could you put the columns in a data frame in alphaetical order?
```{r}
<<<<<<< HEAD
mtcars[order(names(mtcars))]
=======
df3b[order(names(df3b))]
>>>>>>> upstream/master
```
