---
title: "Subsetting"
author: "Alathea"
date: '2014-07-09'
output: 
  html_document:
    keep_md: yes
---

***

### Discussion Notes

***

### Quiz

1. *What does subsetting a vector with: positive integers, negative integers, logical vector, character vector?*

**I guess this means what happens when you subset a vector with one of the above.  A positive integer will select the item at that index and a negative integer will remove the item at that index.  I don't know about the vectors.**

2. *What’s the difference between `[`, `[[` and `$` when applied to a list?*

**Subsetting with `[` will select a list at that location but `[[` selects the item within that list.  `$` works for a named list.**

3. *When should you use `drop = FALSE`?*

**No one knows!**

4. *If x is a matrix, what does `x[] <- 0` do? How is it different to `x <- 0`?*

**Don't know.**

5. *How can you use a named vector to relabel categorical variables?*

**Maybe `names(data_frame) <- names(vector)` or something like that.**

***

### Data Types

#### Atomic Vectors

```{r subsetting_a_vector}
x <- c(1:10)
x[2]
x[c(5,9)]
x[-c(5:10)]
x[c(TRUE,FALSE,TRUE,TRUE,FALSE)]
```

An interesting thing about subsetting with a logical vector is that if the logical vector is shorter than the vector being subsetted, it will act like a pattern and repeat over the entire vector (see above).  I guess this is called recycling.

You can use character vectors to select from a named vector.  I did not know that was possible!

```{r named_vector}
y <- c(1:10)
names(y) <- letters[1:10]
y
```

#### Matrices and Arrays

Arrays are stored in column order.  If you subset with a number, e.g. `array[2]` it will count down through each item of each column.

#### Data Frames

List-type vs. Matrix type subestting:

```{r dataframe_subsetting}
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3])

# There's an important difference if you select a single column:
# matrix subsetting simplifies by default, list subsetting does not.
str(df["x"])

str(df[, "x"])
```

#### S3 vs. S4 Objects

Noooooooooooooooooo!!!

#### Exercises

1. *Fix each of the following common data frame subsetting errors:*

```{r fix_subsetting errors, eval = FALSE}
mtcars[mtcars$cyl = 4, ]
# to select values equal to 4, use "=="
mtcars[mtcars$cyl == 4, ]

mtcars[-1:4, ]
# to remove multiple entries, use a vector
mtcars[-c(1:4), ]

mtcars[mtcars$cyl <= 5]
# also need to select the columns
mtcars[mtcars$cyl <= 5, ]

mtcars[mtcars$cyl == 4 | 6, ]
# you have to do this:
mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]
```

2. *Why does `x <- 1:5; x[NA]` yield five missing values? Hint: why is it different from `x[NA_real_]`?*

**Not sure but something to do with `NA` being logical?**

3. *What does `upper.tri()` return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?*

```{r upper_tri, eval=FALSE}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```

**It returns `TRUE` in the upper triangle of the matrix so subsetting with it returns the values in the upper triangle, column by column.**

4. *Why does `mtcars[1:20]` return an error? How does it differ from the similar `mtcars[1:20, ]`?*

**Cuz you didn't tell it which columns to select.**

5. *Implement your own function that extracts the diagonal entries from a matrix (it should behave like `diag(x)` where `x` is a matrix).*

**Whyyyy????**

```{r diag_function}
x <- outer(1:4, 1:5, FUN = "*")
diag(x)

get_diag <- function(input_matrix)
{
  n <- min(dim(input_matrix))
  select <- vector()
  for(i in 1:n){
    select[i] <- input_matrix[i,i]
  }
  return(select)
}

get_diag(x)
```

6. *What does `df[is.na(df)] <- 0` do? How does it work?*

**It sets `NA` values to 0 by selecting those values (`is.na` generates a data frame with `TRUE` or `FALSE` depending on whether the condition is satisfied)**

***

### Subsetting Operators