---
title: "Subsetting"
author: "Alathea"
date: '2014-07-09'
output: 
  html_document:
    keep_md: yes
    toc: yes
    theme: united
---

***

### Discussion Notes

```{r}
mtcars$cyl == 4 | 6
mtcars$cyl == 4 || 6

# this works
mtcars$cyl == 4 | mtcars$cyl == 6
# or this
mtcars$cyl %in% c(4, 6)
```

```{r eval = FALSE}
# this gives 5 NAs because NA is a logical vector, which gets recycled
x <- 1:5
x[NA]
```

Jenny showed us this:

```{r}
mod <- lm(mpg ~ wt, data = mtcars)
str(mod, max.level = 1, give.attr = FALSE)
```

```{r}
# linearize the output and force back into the original structure
df <- data.frame(x = 1:3, y = 4:6)
df[] <- sample(as.matrix(df))
```

`|` vs. `||`.  The shorter form makes element by element comparisons and the long form compares only the first element of each vector.  Normally the long form is used in `if` statements.

***

### Quiz

1. *What does subsetting a vector with: positive integers, negative integers, logical vector, character vector?*

**I guess this means what happens when you subset a vector with one of the above.  A positive integer will select the item at that index and a negative integer will remove the item at that index.  I don't know about the vectors.**

2. *What’s the difference between `[`, `[[` and `$` when applied to a list?*

**Subsetting with `[` will select a list at that location but `[[` selects the item within that list.  `$` works for a named list.**

3. *When should you use `drop = FALSE`?*

**No one knows!**

4. *If x is a matrix, what does `x[] <- 0` do? How is it different to `x <- 0`?*

**Don't know.**

5. *How can you use a named vector to relabel categorical variables?*

**Maybe `names(data_frame) <- names(vector)` or something like that.**

***

### Data Types

#### Atomic Vectors

```{r subsetting_a_vector}
x <- c(1:10)
x[2]
x[c(5,9)]
x[-c(5:10)]
x[c(TRUE,FALSE,TRUE,TRUE,FALSE)]
```

An interesting thing about subsetting with a logical vector is that if the logical vector is shorter than the vector being subsetted, it will act like a pattern and repeat over the entire vector (see above).  I guess this is called recycling.

You can use character vectors to select from a named vector.  I did not know that was possible!

```{r named_vector}
y <- c(1:10)
names(y) <- letters[1:10]
y
```

#### Matrices and Arrays

Arrays are stored in column order.  If you subset with a number, e.g. `array[2]` it will count down through each item of each column.

#### Data Frames

List-type vs. Matrix type subestting:

```{r dataframe_subsetting}
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3])

# There's an important difference if you select a single column:
# matrix subsetting simplifies by default, list subsetting does not.
str(df["x"])

str(df[, "x"])
```

#### S3 vs. S4 Objects

Noooooooooooooooooo!!!

#### Exercises

1. *Fix each of the following common data frame subsetting errors:*

```{r ex_fix_subsetting errors, eval = FALSE}
mtcars[mtcars$cyl = 4, ]
# to select values equal to 4, use "=="
mtcars[mtcars$cyl == 4, ]

mtcars[-1:4, ]
# to remove multiple entries, use a vector
mtcars[-c(1:4), ]

mtcars[mtcars$cyl <= 5]
# also need to select the columns
mtcars[mtcars$cyl <= 5, ]

mtcars[mtcars$cyl == 4 | 6, ]
# you have to do this:
mtcars[mtcars$cyl == 4 | mtcars$cyl == 6, ]
```

2. *Why does `x <- 1:5; x[NA]` yield five missing values? Hint: why is it different from `x[NA_real_]`?*

**Not sure but something to do with `NA` being logical?**

3. *What does `upper.tri()` return? How does subsetting a matrix with it work? Do we need any additional subsetting rules to describe its behaviour?*

```{r ex_upper_tri, eval=FALSE}
x <- outer(1:5, 1:5, FUN = "*")
x[upper.tri(x)]
```

**It returns `TRUE` in the upper triangle of the matrix so subsetting with it returns the values in the upper triangle, column by column.**

4. *Why does `mtcars[1:20]` return an error? How does it differ from the similar `mtcars[1:20, ]`?*

**Cuz you didn't tell it which columns to select.**

5. *Implement your own function that extracts the diagonal entries from a matrix (it should behave like `diag(x)` where `x` is a matrix).*

**Whyyyy????**

```{r ex_diag_function}
x <- outer(1:4, 1:5, FUN = "*")
diag(x)

get_diag <- function(input_matrix)
{
  n <- min(dim(input_matrix))
  select <- vector()
  for(i in 1:n){
    select[i] <- input_matrix[i,i]
  }
  return(select)
}

get_diag(x)
```

6. *What does `df[is.na(df)] <- 0` do? How does it work?*

**It sets `NA` values to 0 by selecting those values (`is.na` generates a data frame with `TRUE` or `FALSE` depending on whether the condition is satisfied)**

***

### Subsetting Operators

Some subset operations are *simplifying* and drop a bunch of information and others are *preserving* and retain the same structure as the original data.

To access a column with a name contained in another variable, use `[[`. e.g.

```{r get_column_by_variable}
data <- data.frame(a = c(3:5), b = c(10:12), c = c(89:91))
var <- "a"

data$var
data[[var]]
```

The `$` can do partial name matching.  Whoa....dangerous.

```{r partial_name_matching}
data2 <- data.frame(abc = c(3:5), bcd = c(10:12), cde = c(89:91))
data$a
```

#### Exercises

1. *Given a linear model, e.g. `mod <- lm(mpg ~ wt, data = mtcars)`, extract the residual degrees of freedom. Extract the R squared from the model summary (`summary(mod)`)*

```{r ex_model_extraction}
mod <- lm(mpg ~ wt, data = mtcars)
str(mod)

(df_mod <- mod[["df.residual"]])

mod_sum <- summary(mod)
str(mod_sum)

(rsq_mod <- mod_sum[["r.squared"]])
```

### Subsetting and Assignment

Can you combine integer indices with integer NA?

```{r subsetting_w_integers_and_nas}
x <- c(1:5)

# this is working OK.
x[c(1, NA)] <- 0
```

```{r subsetting_w_nothing_preserve_structure}
data(mtcars)
mtcars[] <- lapply(mtcars, as.integer)
str(mtcars[])

mtcars <- lapply(mtcars, as.integer)
str(mtcars)
```

### Applications

Lookup tables:

```{r lookup_table}
students <- data.frame(name = c(LETTERS[1:10]),
                       sex = c("m","f","m","f","f","f","f","m","m","f"),
                       student.number = sample(c(3333:8888), 10))
lookup <- c(m = "Male", f = "Female")

lookup[students$sex]
```

Random samples / bootstrap:

```{r bootstrap}
(x <- data.frame(a = c(1:9), b = letters[1:9]))
x[sample(nrow(x),3),]
```

Ordering:

```{r order}
(y <- x[sample(nrow(x),nrow(x)),])
(y <- y[order(y$a), ])
```

### Exercises

1. *How would you randomly permute the columns of a data frame? (This is an important technique in random forests). Can you simultaneously permute the rows and columns in one step?*

```{r ex_randomize}
(df <- data.frame(a = c(1:4), b = c(5:8), c = c(9:12), d = c(13:16)))

# randomly permute the columns
df[ , sample(ncol(df), ncol(df))]

# randomize columns and rows simultaneously
# although I doubt this is really happening simultaneously
df[sample(nrow(df), nrow(df)), sample(ncol(df), ncol(df))]
```

2. *How would you select a random sample of `m` rows from a data frame? What if the sample had to be contiguous (i.e. with an initial row, a final row, and every row in between)?*

```{r ex_sample}
sample_rows <- function(input_df, sample_size)
{
  input_df[sample(nrow(input_df), sample_size, replace = TRUE), ]
}

sample_rows(df, 3)
sample_rows(df, 6)
```

```{r ex_contig_sample}
sample_contig_rows <- function(input_df)
{
  (row_nums <- sample(nrow(input_df), 2))
  input_df[min(row_nums):max(row_nums), ]
}

sample_contig_rows(df)

# if you want a set number of rows:
sample_contig_rows_2 <- function(input_df, sample_size)
{
  # check that the sample size is smaller than the number of rows
  if(sample_size > nrow(input_df)){
    warning("Sample size too large.  Selecting all rows starting from a random index.")
    return(sample_contig_rows(input_df))
  }
  else if(sample_size == nrow(input_df)){
    warning("You just selected the entire data frame!")
    return(input_df)
  }
  else{
    row_max <- nrow(input_df) - sample_size + 1
    first_row <- sample(1:row_max, 1)
    return(input_df[first_row:(first_row + sample_size - 1), ])
  }
}

sample_contig_rows_2(df, 5)
sample_contig_rows_2(df, 4)
<<<<<<< HEAD
<<<<<<< HEAD
sample_contig_rows_2(df, 2)
=======
sample_contig_rows_2(df, 3)
>>>>>>> upstream/master
=======
sample_contig_rows_2(df, 3)
>>>>>>> upstream/master
```


3. *How could you put the columns in a data frame in alphaetical order?*

```{r ex_alphabetize}
(df3a <- df[ ,sample(ncol(df),ncol(df))])
(df3b <- df3a[ ,order(names(df3a))])
```