#Subsetting

In the *Data frames* section, I don't understand how this subset works, because I don't understand lists:
```
# Like a list:
df <- data.frame(x = 1:3, y = 3:1, z = letters[1:3])
df[c("x", "z")]
#>   x z
#> 1 1 a
#> 2 2 b
#> 3 3 c
```
## Data Types Exercises
 
 1. 
 ```{r}
 mtcars[mtcars$cyl == 4, ]
 mtcars[-c(1:4),]
 mtcars[mtcars$cyl <= 5, ]
 mtcars[mtcars$cyl %in% c(4,6), ]
 ```
 2. `NA` is considered logical in this case, so `R` recycles the index along the length of the vector. Since a missing value in the index always yields a missing value in the output, `NA` is returned 5 times. 
 
 ["Logical. The indexing i should generally have the same length as x. If it is shorter, then its elements will be recycled as discussed in Elementary arithmetic operations. If it is longer, then x is conceptually extended with NAs. The selected values of x are those for which i is TRUE."] (http://cran.r-project.org/doc/manuals/r-release/R-lang.html#Indexing-by-vectors) `NA_real_` is numeric and so is not recycled along the vector.
 
 3. 
 ```{r}
 x <- outer(1:5, 1:5, FUN = "*")
 x[upper.tri(x)]
 ```
 `upper.tri` returns a matrix of logicals the same size of a given matrix with entries `TRUE` in the lower or upper triangle. Subsetting a matrix with it returns the entries of the triangle in column order flattened into a vector. We do not need additional subsetting rules to describe its behavior.
 
 4. `mtcars[1:20]` returns an error because R doesn't know what dimension to subset. `mtcars[1:20,]` specifies both rows and columns (indirectly). THIS IS ALL WRONG. AHHHH. SEE EXAMPLE WITH LISTS IN THE TEXT AND FIGURE OUT EVERYTHING YOU THOUGHT YOU ALREADY KNEW.
 
 5. 
 ```{r}
 mat <- matrix(1:9, nrow = 3, ncol = 3)
 getdiag <- function(yr_mat) {
     s <- dim(yr_mat)
     d <- matrix(c(seq(1,s[1]), seq(1, s[2])), nrow = s[1], ncol = 2)
     yr_mat[d]
 }
 getdiag(mat)
 ```
 
 6.
 ```{r}
 df <- data.frame(x=c(1,2,3), y=c(4,5,6))
 is.na(df$x) <- 1 
 df[is.na(df)] <- 0 
 ```
 `df[is.na(df)] <- 0` sets all `NA`s in a data frame to `0`. `is.na(df)` creates a data frame of logicals. `[` selects the `TRUE` values and `<- 0` assigns them to 0.
 
 
 
 