---
title: "Expressions"
author: "Alathea DL"
date: '2014-12-02'
output:
  html_document:
    keep_md: yes
    theme: united
    toc: yes
---

# Exercises

## Structure of expressions

### There’s no existing base function that checks if an element is a valid component of an expression (i.e., it’s a constant, name, call, or pairlist). Implement one by guessing the names of the “is” functions for calls, names, and pairlists.

### pryr::ast() uses non-standard evaluation. What’s its escape hatch to standard evaluation?

### What does the call tree of an if statement with multiple else conditions look like?

### Compare ast(x + y %+% z) to ast(x ^ y %+% z). What do they tell you about the precedence of custom infix functions?

### Why can’t an expression contain an atomic vector of length greater than one? Which one of the six types of atomic vector can’t appear in an expression? Why?

## Names

### You can use formals() to both get and set the arguments of a function. Use formals() to modify the following function so that the default value of x is missing and y is 10.

```{r names_1, eval = FALSE}
g <- function(x = 20, y) {
      x + y
    }
```

### Write an equivalent to get() using as.name() and eval(). Write an equivalent to assign() using as.name(), substitute(), and eval(). (Don’t worry about the multiple ways of choosing an environment; assume that the user supplies it explicitly.)

## Calls

### The following two calls look the same, but are actually different. What’s the difference? Which one should you prefer?

```{r calls_1, eval = FALSE}
(a <- call("mean", 1:10))
#> mean(1:10)
(b <- call("mean", quote(1:10)))
#> mean(1:10)
identical(a, b)
#> [1] FALSE

class(a)
class(b)

eval(a)
eval(b)
```

`quote` returns `1:10` as an expression

### Implement a pure R version of do.call().

```{r my_do_call}
my_do_call <- function(function_name, ...){
  # get the call to my_do_call
  .call <- match.call()
  
  # remove the "my_do_call" function from the call
  my_call <- as.list(.call[-1])
  
  # evaluate the new call
  eval(as.call(my_call))
}

my_do_call(mean, 1:10)
```


### Concatenating a call and an expression with c() creates a list. Implement concat() so that the following code works to combine a call and an additional argument.

```{r concat, eval = FALSE}
concat(quote(f), a = 1, b = quote(mean(a)))
#> f(a = 1, b = mean(a))
```

### Since list()s don’t belong in expressions, we could create a more convenient call constructor that automatically combines lists into the arguments. Implement make_call() so that the following code works.

```{r make_call, eval = FALSE}
make_call(quote(mean), list(quote(x), na.rm = TRUE))
#> mean(x, na.rm = TRUE)
make_call(quote(mean), quote(x), na.rm = TRUE)
#> mean(x, na.rm = TRUE)
```

### How does mode<- work? How does it use call()?

### Read the source for pryr::standardise_call(). How does it work? Why is is.primitive() needed?

### standardise_call() doesn’t work so well for the following calls. Why?

```{r standardise_call, eval = FALSE}
standardise_call(quote(mean(1:10, na.rm = TRUE)))
#> mean(x = 1:10, na.rm = TRUE)
standardise_call(quote(mean(n = T, 1:10)))
#> mean(x = 1:10, n = T)
standardise_call(quote(mean(x = 1:10, , TRUE)))
#> mean(x = 1:10, , TRUE)
```

### Read the documentation for pryr::modify_call(). How do you think it works? Read the source code.

### Use ast() and experimentation to figure out the three arguments in an if() call. Which components are required? What are the arguments to the for() and while() calls?


## Capturing the current call

## Pairlists

## Parsing and deparsing

## Walking the call tree with recursive functions

***
***

# Notes

## Structure of expressions

* use `quote()` to capture an expression

```{r structure_of_expressions}
require(pryr)

# Four components of expressions
# 1: constants
ast("a")
ast(1)

#2: names / symbols
ast(x)

#3: calls (every operation is a function call)
ast(f())
ast(f(1,2))
ast(a + b)

```

## Names

* Names are also called symbols. `as.symbol()` = `as.name()`

## Calls

* calls are similar to lists
* calls have `length`, `[[` and `[`
* first element is the function
* remaining elements are the arguments
* calls can be modified using `[]` and `<-`
* use `call` or `as.call` to create your own call

## Capturing the current call

## Pairlists

## Parsing and deparsing

## Walking the call tree with recursive functions