---
title: "Jenny's reading of 01_DataStructures"
author: "Jenny Bryan"
date: "7 July, 2014"
output:
  html_document:
    toc: yes
---

Week 01 2014-07-03 we read [Data Structures](http://adv-r.had.co.nz/Data-structures.html).

## Taking the quiz

*I made myself enter these answers before reading the chapter, most especially before reading the answers. I annotated/corrected my original answers as I read on.*

#### What are the three properties of a vector, other than its contents?

  * length, "flavor", names
  
*"length" was correct. My "flavor" is more precisely captured by "type" via `typeof()`. And "names" was too specific; the more general "attributes" was more correct.*

#### What are the four common types of atomic vectors?

  * character, logical, numeric, factor

*I was right re: logical and character. But I should have split "numeric" out into integer and double. I should not have mentioned factor which is, I guess, of type integer.*

#### What are the two rare types?

  * I don't know.
  
*Complex and raw. Now I know.*

#### What are attributes? How do you get them and set them?

  * I think of them as an arbitrary way to add meta data to objects. I think names are an example of a common attribute. Use `attributes()` to query and set them.
  
*My answer was pretty good. Refinement: `attributes()` accesses all attributes at once as a list, where as `attr()` is for individual access.*

#### How is a list different from an atomic vector? How is a matrix different from a data frame?

  * Elements of a list don't need to have the same "flavor".
  * A matrix must be homogeneous, i.e. all elements have same "flavor". A data frame is a special list: each element is an atomic vector, each vector has same length, but the vectors don't necessarily have same "flavor".

#### Can you have a list that is a matrix? Can a data frame have a column that is a matrix?

  * Yes a list can be a matrix. At least, I know that a list can have dimensions. Does that make it a matrix?
  * And, yes, I think a data frame can have a column that is a matrix. But it fills me with dread.

## Vectors

Good to know:

  * `is.vector()` does not test if an object is a vector. Instead it returns `TRUE` only if the object is a vector with no attributes apart from names. Use `is.atomic(x) || is.list(x)` to test if an object is actually a vector.
  * `is.double()` tests for doubles whereas `is.numeric()` "tests for the numberliness of a vector and returns `TRUE` for both integer and double vectors".
  * `c()` flattens but `list()` does not

### Working the vectors exercises

#### What are the six types of atomic vector?

  * logical, character, integer, double, complex, raw

#### How does a list differ from an atomic vector?

  * List elements don't necessarily have same type, whereas atomic vector elements do.

#### What makes `is.vector()` and `is.numeric()` fundamentally different to `is.list()` and `is.character()`?

  * Although `is.vector()`'s suggests it would test if an object is a vector, it does not. It seems to test more whether it is just a vector -- of the most basic kind. It tests if "the object is a vector with no attributes apart from names". `is.atomic()` is more suitable for testing if an object is a vector. `is.list()` tests whether an object is truly a list. `is.numeric()` tests for "numberliness" and will return `TRUE` for integer or double atomic vectors. This is different from `is.character()` which is testing for one of the basic types.*

#### Test your knowledge of vector coercion rules by predicting the output of the following uses of `c()`:

  * `c(1, FALSE)` integer vector of 1 and 0 __close but wrong__ type is actually double
  * `c("a", 1)` character vector of "a" and "1" __correct__
  * `c(list(1), "a")` list; element 1 is a list containing integer vector 1 and element 2 is a list containing character vector "a" __correct__
  * `c(TRUE, 1L)` integer vector of 1 and 1 __correct__
  
#### Why do you need to use `unlist()` to convert a list to an atomic vector? Why doesn’t `as.vector()` work?

  * `as.vector()` won't work because I think lists are already vectors, just not atomic vectors. I know I can create a list using `vector(mode = "list")`, which is why I think applying `as.vector()` to a list object won't actually convert it.

#### Why is `1 == "1"` true?

  * To test for equality, I'm guessing both sides are coerced to a sensible type which, in this case, will be character. And so we'll get `TRUE`.

#### Why is `-1 < FALSE` true?

  * To evaluate this inequality, again there will be coercion and the `FALSE` will become zero, which is greater than -1.

#### Why is `"one" < 2` false?

  * I have no idea.

#### Why is the default missing value, NA, a logical vector? What’s special about logical vectors? (Hint: think about c(FALSE, NA_character_).)

  * I think of logicals as the most finicky type and the most vulnerable to coercion. So if you make `NA`s logical, coercion will kick in when and if necessary and you'll eventually get to the appropriate type of `NA`.

## Attributes

Good to know:

  * `structure()` returns a new object with attributes as you set them in the call, e.g. `structure(1:10, my_attribute = "This is a vector")`
  * attributes are easy to lose, e.g. when you index into or compute on an object
  * luckily, the 3 most important attributes are not easily lost and they are
    - names
    - dimensions
    - class
  * name a vector at creation, later via `names()` or via copy and `setNames()`
  * de-name in place via `names(x) <- NULL` or via copy and `unname()`
  
### Working the attributes exercises

#### An early draft used this code to illustrate `structure()`:

```
structure(1:5, comment = "my attribute")
#> [1] 1 2 3 4 5
```

But when you print that object you don’t see the comment attribute. Why? Is the attribute missing, or is there something else special about it? (Hint: try using help).

  * Apparently a comment attribute is a "thing" that has been anticipated in R; I guess we are expected to use it to annotate data frames and model fits. Who knew? Anyway, according to [the help](http://stat.ethz.ch/R-manual/R-devel/library/base/html/comment.html), "contrary to other attributes, the comment is not printed (by `print` or `print.default`)."

#### What happens to a factor when you modify its levels?

  * All hell breaks loose and you will wish you were dead.

```
f1 <- factor(letters)
levels(f1) <- rev(levels(f1))
```

  * When we first create `f1`, it will print as `a b c ...` and have levels `a b c ...`. The above resetting of levels will not cause an error because all the original levels appear -- they're just in a new order. __However__ even I, veteran of many factor battles, got caught out here somewhat. I knew that the levels of `f1` would become `z y x ...` but I was surprised to see that `f1` itself now printed as `z y x ...`. I'm going to explore this further below. Here's the above code in a proper runnable R chunk:

```{r}
f1 <- factor(letters)
f1
str(f1)
levels(f1) <- rev(levels(f1))
f1
str(f1)
```

  * So we see that the underlying integer codes never change. But resetting the levels as we do above makes `1` become associated with `z` instead of `a` and so on. I will make a separate R markdown chunk to delve more deeply into changing a factor's levels this way because more evil surprises lurk.

#### What does this code do? How do `f2` and `f3` differ from `f1`?

```
f2 <- rev(factor(letters))
f3 <- factor(letters, levels = rev(letters))
```

  * `f2` will be a factor. It will print `z x y ...` but have levels `a b c ...`. `f3` will be a factor that prints `a b c ...` with levels `z y x ...`. Here's the above code in a proper runnable R chunk:

```{r}
f2 <- rev(factor(letters))
f2
str(f2)
f3 <- factor(letters, levels = rev(letters))
f3
str(f3)
```

## Matrices and arrays

No notes.

### Working the matrices and arrays exercises

#### What does `dim()` return when applied to a vector?

  * `NULL`

```{r}
dim(letters)
```

#### If `is.matrix(x)` is `TRUE`, what will `is.array(x)` return?

  * `TRUE`

```{r}
(yo <- matrix(letters[1:12], nrow = 3))
is.matrix(yo)
is.array(yo)
```

#### How would you describe the following three objects? What makes them different to `1:5`?

```
x1 <- array(1:5, c(1, 1, 5))
x2 <- array(1:5, c(1, 5, 1))
x3 <- array(1:5, c(5, 1, 1))
```

  * `x1`, `x2`, and `x3` are all multi-dimensional (three-dimensional, actually) arrays, whereas `1:5` is an atomic vector. Granted, they each have only one dimension whose extent is greater than one, but that is still a totally different beast from an atomic vector. I think of them like so: `x1` has 1 row, 1 column and 5 slices. `x2` has 1 row, 5 columns and 1 slice. `x3` has 5 rows, 1 column and 1 slice.

## Data frames

No notes.

### Working the data frames exercises

#### What attributes does a data frame possess?

  * `class`, `dim` and `names` at the very last. Probably `rownames` too, or maybe I should just generalize to `dimnames`. *see the next chunk for the actual answer*
  
```{r}
foo <- data.frame(a = 1:5, b = letters[1:5])
attributes(foo)
```

#### What does `as.matrix()` do when applied to a data frame with columns of different types?

  * I predict coercion happens and it's likely everything ends up as character. *see the next chunk for the actual answer*
  
```{r}
foo <- data.frame(a = 1:5, b = letters[1:5])
str(as.matrix(foo))
```

#### Can you have a data frame with 0 rows? What about 0 columns?

  * Yes, you can definitely have a data frame with 0 rows. Not sure about the 0 columns. Probably not.
