---
title: "Andrew_FOs"
author: "Andrew MacDonald"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: yes
    keep_md: TRUE
---

1. 

```{r}
deparse(substitute(a + b + c + d + e + f + g + h + i + j + k + l + m +
  n + o + p + q + r + s + t + u + v + w + x + y + z))
```

Write a wrapper so that it is always a string:

```{r}
onedeparse <- function(expr){
  paste(deparse(expr), collapse = "")
}

onedeparse(substitute(a + b + c + d + e + f + g + h + i + j + k + l + m +
  n + o + p + q + r + s + t + u + v + w + x + y + z))
```
(does that miss the point?)

2. 
```{r error=TRUE}
as.Date(TRUE)
as.Date.default
```
uses deparse(substitute(x)) in order to make an informative error message that contains the thing you tried to make into a logical variable.

3. 
```{r}
pairwise.t.test
```
used to construct the string stored as `DNAME`, which is used to give informative output featuring the names of the input variables

```{r}
test <- data.frame(fac = gl(2, k = 5), dat = rnorm(10))
pairwise.t.test(test$dat, test$fac)
```

```{r}
test2 <- data.frame(factorthatwehopehassomeexplanatorypowerbutthenagainitmightnot = gl(2, k = 5), dat = rnorm(10))
pairwise.t.test(test2$factorthatwehopehassomeexplanatorypowerbutthenagainitmightnot, test2$fac)
```

I thought that these would return f(whatever_was_in_there) but it didn't, it returned X every time. But Why?  Because 
```{r}
deparse(substitute(x))
```
is the only thing that gets evalutated?  x is never evaluated?

```{r error = TRUE}
f <- function(a) substitute(a) 

g <- function(x) deparse(f(x))
g(1:10)
g(x)
g(x + y ^ 2 / z + exp(a * sin(b)))
```

oh yeah. bad idea!

This will be 4
```{r}
eval(quote(eval(quote(eval(quote(2 + 2))))))
```

also 4 (`eval(4)`)

```{r}
eval(eval(quote(eval(quote(eval(quote(2 + 2)))))))
```

"4"
```{r}
quote(eval(quote(eval(quote(eval(quote(2 + 2)))))))
```
haha NO. it doesn't go from the inside out!

out of curiosity...
```{r}
library(magrittr)

1:3 %>%
  mean %>%
  quote
```

## bug in subset

```{r}
subset2 <- function(x, condition) {
  res <- x
  condition_call <- substitute(condition)
  r <- eval(condition_call, x)
  x[r, , drop = FALSE]
  }

sample_df2 <- data.frame(x = 1:10)
subset2(sample_df2, x > 8)
```

# quote instead of substitute

```{r error = TRUE}
subset2 <- function(x, condition) {
  res <- x
  condition_call <- quote(condition)
  r <- eval(condition_call, x)
  res <- x[r, ]
  res <- as.data.frame(res)
  names(res) <- names(x)
  res
}

sample_df2 <- data.frame(x = 1:10)
subset2(sample_df2, x > 8)
```

## how does this work?
```
select <- function(df, vars) {
  vars <- substitute(vars)
  var_pos <- setNames(as.list(seq_along(df)), names(df))
  pos <- eval(vars, var_pos)
  df[, pos, drop = FALSE]
}
select(mtcars, -cyl)
```
makes a named vector of column positions, and then uses `eval` to 

# `evalq`

looks like evalq, unsurprisingly, is simply a shortcut for`eval(quote(expr), ...)`
```{r, error=TRUE}
eval(quote(2 +2))
evalq(2 + 2)

truth <- list(fact = "science is cool", fact2 = "but its hard to find work")
evalq(truth)
# evalq(fact)
evalq(fact, truth)
```

****

# `arrange`

```{r}
plyr::arrange
```

# transform
```{r}
transform.data.frame
```

evaluates a list of the column names (actually, the expressions on column names) in the original data frame.


```{r}
plyr::mutate
```

`mutate` employs a for loop wherein each expression is evaluated in the environment of the data frame. transform does it all up front in one line.

# `with`

`with` was/is one of my favourite functions when I was first learning.  it lets you refer to variables in a dataframe or list by name, without attaching:

```{r}
df <- data.frame(a = 1, b = 2)
with(df, a + b)
```

But how does it work?

```{r}
with.default
```
 Very simply, by capturing the expression and evaluating it with the data as an environment! setting`enclos` to the parent allows global variables to mix with data-frame variables:
 
 
```{r}
foo <- 7
with(df, a + foo)
```

On the other hand, I never got into the habit of using `within`.  I understand that it works like transform:

```{r}
tl <- data.frame(a = rnorm(20), b = rpois(20, 10))

tl <- within(tl, new <- sum(a, b))
tl
```

er. It doesn't!

```{r}
within.data.frame
```

Wow. very hard to read! but because we are replacing part of an object (rather than simply returning the result of evaluating and expression) its bound to be that way.

## Base escape hatches
```{r}
rm(tl)
rm("foo")
```
`rm` allows a vetor of characters

`library` is strange -- it accepts both quoted unquoted, unless `character.only = TRUE`, then only character.

## cf different function



