---
title: "Exercises for 'Non-standard Evaluation'"
author: "Davor Cubranic"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:
  html_document:
    toc: no
    keep_md: TRUE
---

### Capturing expressions

1.  One important feature of `deparse()` to be aware of when programming is that 
    it can return multiple strings if the input is too long. For example, the 
    following call produces a vector of length two:

    ```{r, eval = FALSE}
    g(a + b + c + d + e + f + g + h + i + j + k + l + m +
      n + o + p + q + r + s + t + u + v + w + x + y + z)
    ```

    Why does this happen? Carefully read the documentation. Can you write a
    wrapper around `deparse()` so that it always returns a single
    string?

    Because `deparse` breaks lines to have lenght of (at least)
    `width.cutoff` bytes. We can create a wrapper that simply calls
    `paste(..., collapse = '')`.

1.  Why does `as.Date.default()` use `substitute()` and `deparse()`?

    To construct the error message if it can't convert the argument.
    
    Why does `pairwise.t.test()` use them? Read the source code.

    To create the label of the table

1.  `pairwise.t.test()` assumes that `deparse()` always returns a length one 
    character vector. Can you construct an input that violates this expectation? 
    What happens?

    ```{r}
    attach(airquality)
    Month <- factor(Month, labels = month.abb[5:9])
    pairwise.t.test(Ozone+Ozone+Ozone+Ozone+Ozone+Ozone+Ozone+Ozone+Ozone+Ozone+Ozone+Ozone+Ozone+Ozone, Month)
    detach()
    ```

    The table label is created by pasting the result of two
    `deparse`s, and because the first one ('Ozone+...') is longer than
    one element, the second ('Month') gets recycled.
    
1.  `f()`, defined above, just calls `substitute()`. Why can't we use it
    to define `g()`? In other words, what will the following code return?
    First make a prediction. Then run the code and think about the results.

    ```{r, eval = FALSE}
    f <- function(x) substitute(x)
    g <- function(x) deparse(substitute(x))
    g2 <- function(x) deparse(f(x))
    g2(1:10)
    g2(x)
    g2(x + y ^ 2 / z + exp(a * sin(b)))
    ```

    It always returns 'x', from the argument to `g2`.


### Non-standard evaluation in `subset`

1.  Predict the results of the following lines of code:

    ```{r, eval = FALSE}
    eval(quote(eval(quote(eval(quote(2 + 2))))))
    eval(eval(quote(eval(quote(eval(quote(2 + 2)))))))
    quote(eval(quote(eval(quote(eval(quote(2 + 2)))))))
    ```

    Prediction: 4, 4, expression 4
    Correct: 4, 4, `eval(quote(eval(quote(eval(quote(2 + 2))))))`
    
1.  `subset2()` has a bug if you use it with a single column data frame.
    What should the following code return? How can you modify `subset2()`
    so it returns the correct type of object?

    Add `drop=FALSE` when selecting the rows to return:
    ```{r}
    subset2 <- function(x, condition) {
      condition_call <- substitute(condition)
      r <- eval(condition_call, x)
      x[r, , drop=FALSE]
    }
    sample_df2 <- data.frame(x = 1:10)
    subset2(sample_df2, x > 8)
    ```

1.  The real subset function (`subset.data.frame()`) removes missing
    values in the condition. Modify `subset2()` to do the same: drop the 
    offending rows.

    ```{r}
    subset2(data.frame(x=c(1, 2, NA, 44)), x < 4)
    subset3 <- function(x, condition) {
      condition_call <- substitute(condition)
      r <- eval(condition_call, x)
      x[!is.na(r) & r, , drop=FALSE]
    }
    subset3(data.frame(x=c(1, 2, NA, 44)), x < 4)
    ```

1.  What happens if you use `quote()` instead of `substitute()` inside of
    `subset2()`?

    ```{r}
    sample_df <- data.frame(a = 1:5, b = 5:1, c = c(5, 3, 1, 4, 1))
    subset_quote <- function(x, condition) {
      condition_call <- quote(condition)
      r <- eval(condition_call, x)
      x[!is.na(r) & r, , drop=FALSE]
    }
    subset_quote(sample_df, a >= 4)
    ```

    Using `quote` will make 'condition_call' equal to symbol 'condition', so
    that `eval`-ing it will cause it to be looked-up in
    `subset_quote`'s caller, where it won't be found, causing an
    error.
    
1.  The second argument in `subset()` allows you to select variables. It
    treats variable names as if they were positions. This allows you to do 
    things like `subset(mtcars, , -cyl)` to drop the cylinder variable, or
    `subset(mtcars, , disp:drat)` to select all the variables between `disp`
    and `drat`. How does this work? I've made this easier to understand by
    extracting it out into its own function.

    ```{r, eval = FALSE}
    select <- function(df, vars) {
      vars <- substitute(vars)
      var_pos <- setNames(as.list(seq_along(df)), names(df))
      pos <- eval(vars, var_pos)
      df[, pos, drop = FALSE]
    }
    select(mtcars, -cyl)
    ```

    It first creates an index of variable positions, as 'var_pos',
    with variable names as element names and position as values. It
    then evaluates the 'vars' expression (e.g., `-cyl`) using the
    index as the environment, so that variable names will be replaced
    by their position, which can then be used as the column index into
    the original data frame.
    
1.  What does `evalq()` do? Use it to reduce the amount of typing for the
    examples above that use both `eval()` and `quote()`.

    It is equivalent to `eval(quote(expr), ...)`, but we rarely used
    `eval(quote(...))`, but rather `eval(substitute(...))`.


### Scoping issues

1.  `plyr::arrange()` works similarly to `subset()`, but instead of selecting
    rows, it reorders them. How does it work? What does
    `substitute(order(...))` do? Create a function that does only that
    and experiment with it.

    `substitute(order(...))` returns an expression in which `...` is
    replaced by the quoted arguments that were passed as part of
    `...`. For instance, if the call were `substitute(mtcars, cyl,
    mpg)`, then `...` is '`cyl, mpg`', and the whole expression is
    '`order(cyl, mpg)`'. When this is evaluated in the environment of
    the data frame, it will return the new indexing of the rows of the
    data frame which will put them in order of columns in 'cyl' and
    'mpg' (i.e., whatever was in `...`).

1.  What does `transform()` do? Read the documentation. How does it work?
    Read the source code for `transform.data.frame()`. What does
    `substitute(list(...))` do?

    When given a data frame some expressions as named arguments, it
    will evaluate the expressions in the environment of the data frame
    and use their result to replace the variables in the data frame if
    the argument's name matches the variable, or append them to the
    data frame as a new variable.

1.  `plyr::mutate()` is similar to `transform()` but it applies the
    transformations sequentially so that transformation can refer to columns
    that were just created:

    ```{r, eval = FALSE}
    df <- data.frame(x = 1:5)
    transform(df, x2 = x * x, x3 = x2 * x)
    plyr::mutate(df, x2 = x * x, x3 = x2 * x)
    ```

    How does mutate work? What's the key difference between `mutate()` and
    `transform()`?

    `mutate` loops through the named arguments, evaluates them in the
    environment of the data frame, and puts the result into the data
    frame before proceeding to the next argument, so that it can refer
    to those that preceeded it.

1.  What does `with()` do? How does it work? Read the source code for
    `with.default()`. What does `within()` do? How does it work? Read the
    source code for `within.data.frame()`. Why is the code so much more
    complex than `with()`?

    Given a data frame and an expression, `with` simply does a
    `eval(substitute(expr), df)`, evaluating the expression in the
    environment of the data frame. On the other hand, `within` checks
    whether the evaluating the expression caused any changes in the
    environment, and if so, makes the same changes to the (copy of
    the) data frame and returns it.
    
### Calling from another function

1.  The following R functions all use NSE. For each, describe how it uses NSE,
    and read the documentation to determine its escape hatch.
    * `rm()`
    * `library()` and `require()`
    * `substitute()`
    * `data()`
    * `data.frame()`

1.  Base functions `match.fun()`, `page()`, and `ls()` all try to
    automatically determine whether you want standard or non-standard
    evaluation. Each uses a different approach. Figure out the essence
    of each approach then compare and contrast.

1.  Add an escape hatch to `plyr::mutate()` by splitting it into two functions.
    One function should capture the unevaluated inputs. The other should take a 
    data frame and list of expressions and perform the computation.

1.  What's the escape hatch for `ggplot2::aes()`? What about `plyr::()`?
    What do they have in common? What are the advantages and disadvantages
    of their differences?

1.  The version of `subset2_q()` I presented is a simplification of real
    code. Why is the following version better?

    ```{r}
    subset2_q <- function(x, cond, env = parent.frame()) {
      r <- eval(cond, x, env)
      x[r, ]
    }
    ```

    Rewrite `subset2()` and `subscramble()` to use this improved version.


### NSE and escape hatches

1.  Use `subs()` to convert the LHS to the RHS for each of the following pairs:
    * `a + b + c` -> `a * b * c`
    * `f(g(a, b), c)` -> `(a + b) * c`
    * `f(a < b, c, d)` -> `if (a < b) c else d`

2.  For each of the following pairs of expressions, describe why you can't
    use `subs()` to convert one to the other.
    * `a + b + c` -> `a + b * c`
    * `f(a, b)` -> `f(a, b, c)`
    * `f(a, b, c)` -> `f(a, b)`

3.  How does `pryr::named_dots()` work? Read the source.


### Downsides of NSE

1.  What does the following function do? What's the escape hatch?
    Do you think that this is an appropriate use of NSE?

    ```{r}
    nl <- function(...) {
      dots <- named_dots(...)
      lapply(dots, eval, parent.frame())
    }
    ```

2.  Instead of relying on promises, you can use formulas created with `~`
    to explicitly capture an expression and its environment. What are the
    advantages and disadvantages of making quoting explicit? How does it
    impact referential transparency?

3.  Read the standard non-standard evaluation rules found at
    <http://developer.r-project.org/nonstandard-eval.pdf>.

