---
title: "Environments-Andrew"
author: "Andrew MacDonald"
date: '2014-07-24'
output: 
  html_document:
    keep_md: yes
    toc: yes
---

```{r setup, include = FALSE, cache = FALSE}
knitr::opts_chunk$set(error = TRUE)
```


# Quiz

1.  List three ways that an environment is different to a list.

lists are ordered, environments are not
lists are objects
um, copy-on-modify?


1.  What is the parent of the global environment? What is the only 
    environment that doesn't have a parent?
The empty environment, which has no parent    

1.  What is the enclosing environment of a function? Why is it 
    important?

A function's enclosing environment is the environment in which it was created.  
It contains all the objects that were present when the function was defined.

1.  How do you determine the environment from which a function was called?

uhm, idk. `environment`?

1.  How are `<-` and `<<-` different?

`<-` assigns an object to a name in the current environment.  `<<-` assigns in the parent as well.  Additionally, the latter is evil.

## Exercises
### List three ways in which an environment differs from a list.
* NO order
* everything must have a unique name
* environments have parents

### If you don’t supply an explicit environment, where do ls() and rm() look? Where does <- make bindings?

I'm betting that this is in the present environment.  Is there a way to test that?

### Using parent.env() and a loop (or a recursive function), verify that the ancestors of globalenv() include baseenv() and emptyenv(). Use the same basic idea to implement your own version of search().

```{r}
library(magrittr)
parent.env(parent.env(environment()))

environment() %>% parent.env %>% parent.env

the_env <- globalenv()

while(!identical(the_env,emptyenv())) {
  print(the_env)
  the_env <- parent.env(the_env)
  }

parent_finder <- function(e = globalenv()) {
  if(identical(e, emptyenv())) {
    print(e)
    } else {
      print(e)
      parent_finder(parent.env(e))
      }
    }

parent_finder()

parent_list <- function(e = globalenv()) {
  envlist <- list()
  if(identical(e, emptyenv())) {
    e
    } else {
      envlist <- c(envlist,e)
      parent_list(parent.env(e))
      }
  envlist
    }

foo <- parent_list()

```

## Exercises 2

### Modify where() to find all environments that contain a binding for name.

```{r}

where_all <- function(name, env = parent.frame()) {
  env_list <- list()
  if (identical(env, emptyenv())) {
    # base case
    stop("Can't find ", name, call. = FALSE)
  } else if (exists(name, envir = env, inherits = FALSE)) {
    # success case
    env_list <- c(env_list,env)
  } else {
    # recursive case
    where_all(name, env = parent.env(env))
  }
  env_list
}

library(magrittr)
library(dplyr)

where_all("%>%")
```

Why didn't that work??


### Write your own version of get() using a function written in the style of where().

```{r}
where_get <- function(name, env = parent.frame()) {
  if (identical(env, emptyenv())) {
    # Base case
    stop("Can't find ", name, call. = FALSE)
    
  } else if (exists(name, envir = env, inherits = FALSE)) {
    # Success case
    eval(as.name(name),env)
    
  } else {
    # Recursive case
    where_get(name, parent.env(env))
    
  }
}

x <- 4
e <- new.env()
e$y <- 5

where_get("y",env = e)
where_get("x",env = e)
```

### Write a function called fget() that finds only function objects
 It should have two arguments, name and env, and should obey the regular scoping rules for functions: if there’s an object with a matching name that’s not a function, look in the parent. For an added challenge, also add an inherits argument which controls whether the function recurses up the parents or only looks in one environment.
 
```{r}
fget <- function(name, env = parent.frame(), inherits = TRUE) {
  if (identical(env, emptyenv())) {
    # Base case
    stop("Can't find ", name, call. = FALSE)
    
    } else if (exists(name, envir = env, inherits = FALSE) & 
                 (eval(as.name(name), envir = env) %>% is.function) ) {
      # Success case
      eval(as.name(name),env)
      
      } else if(inherits) {
        # Recursive case
        fget(name, parent.env(env))
        
        } else {
          stop("Não tem não")
          }
  }

e$mean <- 100
fget("mean")

fget("mean",e)

fget("mean",e, inherits = FALSE)
```

### Write your own version of exists(inherits = FALSE)

```{r}
andrews_exists <- function(name, env = parent.frame()) {
  if (identical(env, emptyenv())) {
    # Base case
    stop("Can't find ", name, call. = FALSE)
    
    } else {
      # Success case
      any(name %in% ls(env))
      }
  }

andrews_exists("y",e)

andrews_exists("y")
```

Recursive??

```{r}


```

