---
title: "Environments"
author: "Alathea"
date: '2014-07-21'
output: 
  html_document:
    keep_md: yes
    toc: yes
---

***

## Exercises

### List three ways in which an environment differs from a list.

* The environment has a parent
* Objects in an environment have no order
* Every object in an environment has to have a unique name

### If you don’t supply an explicit environment, where do `ls()` and `rm()` look? Where does `<-` make bindings?

They look in the current environment (usually `globalenv()`) and that is also where `<-` makes bindings.  Within a functions, the bindings are made in the function's environment.

### Using `parent.env()` and a loop (or a recursive function), verify that the ancestors of `globalenv()` include `baseenv()` and `emptyenv()`. Use the same basic idea to implement your own version of `search()`.

```{r ancestors, message = FALSE}
ancestors <- function(environ = globalenv())
{
  p <- parent.env(environ)
  print(environmentName(p))
  
  if(identical(p, emptyenv())) return(invisible())
  else ancestors(p)
}

ancestors()
```

### Modify `where()` to find all environments that contain a binding for `name`.

I'm not entirely sure I understand the question, but:

```{r where_name, message = FALSE}
my_where <- function(env = parent.frame()) {
  if (identical(env, emptyenv())) {
    # Base case
    stop("Can't find 'name'", call. = FALSE)
  } else if (exists("name", envir = env, inherits = FALSE)) {
    # Success case
    env 
  } else {
    # Recursive case
    my_where(parent.env(env))
  }
}

name <- "Alathea"
my_where()
```

### Write your own version of `get()` using a function written in the style of `where()`.

```{r get, message = FALSE}
my_get <- function(obj_name, env = parent.frame(), recursive = TRUE)
{
  if (exists(obj_name, envir = env, inherits = FALSE)) {
    eval(parse(text = obj_name))
  } else if (!recursive) {
    stop("Can't find ", obj_name, " in ", environmentName(env), call. = FALSE)
  } else if (identical(env, emptyenv())) {
    stop(obj_name, " does not exist", call. = FALSE)
  } else my_get(obj_name, parent.env(env))
}

name <- "Alathea"

get("name")
my_get("name")
```

### Write a function called `fget()` that finds only function objects. It should have two arguments, `name` and `env`, and should obey the regular scoping rules for functions: if there’s an object with a matching name that’s not a function, look in the parent. For an added challenge, also add an inherits argument which controls whether the function recurses up the parents or only looks in one environment.

```{r fget, message = FALSE}
fget <- function(name, env, recursive = FALSE){
  g <- get(name)
  if(!is.function(g)) stop("The 'name' argument must specify a function.")
  return(g)
}

fget("sum")
```

### Write your own version of `exists(inherits = FALSE)` (Hint: use `ls()`). Write a recursive version that behaves like `exists(inherits = TRUE)`s.

```{r my_exists, message = FALSE}
my_exists_local <- function(name)
{
  if(!is.character(name)) stop("'name' must be a character string.  Please try again.")
  
  name_list <- ls(envir = parent.frame())
  
  if(any(name_list == name)) {
    print(paste("Yes it exists and the value is:", name, sep = " "))
  } else {
    print("No it doesn't exist.")
  }
}

my_exists_global <- function(name, environ = parent.frame())
{
  if(!is.character(name)) stop("'name' must be a character string.  Please try again.")
  
  name_list <- ls(envir = environ)
  if(identical(environ, emptyenv())) {
    print("No it doesnt exist.")
  } else if(any(name_list == name)) {
    print(paste("Yes it exists and the value is:", get(name), sep = " "))
  } else {
    environ = parent.env(environ)
    my_exists_global(name, environ)
  }
}

assign("do_i_exist", "yes i do exist.", envir = baseenv())

my_exists_local("do_i_exist")
my_exists_global("do_i_exist")
```

### Write an enhanced version of `str()` that provides more information about functions. Show where the function was found and what environment it was defined in.

```{r enhanced_str, message = FALSE}
str_plus <- function(fun)
{
  if(!is.function(fun)) stop("Input must be a function.")
 
  require(pryr)
  
  output <- list(structure = str(fun), 
                 binding_env = where(as.character(substitute(fun))), 
                 enclosing_env = environment(fun)
                 )
  output  
}

str_plus(str_plus)
```

***

## Reading Notes

### Quiz

#### List three ways that an environment is different to a list.

from Hadley:

> * Every object in an environment has a unique name.
> * The objects in an environment are not ordered (i.e. it doesn’t make sense to ask what the first object in an environment is).
> * An environment has a parent.
> * Environments have reference semantics.

#### What is the parent of the global environment? What is the only environment that doesn’t have a parent?

The parent of the global environment is the package that was attached last.  The **empty** environment (`emptyenv`) has no parent.

#### What is the enclosing environment of a function? Why is it important?

#### How do you determine the environment from which a function was called?

`parent.frame()`

#### How are `<-` and `<<-` different?

`<-` sets a value in the current environment. `<<-` sets the value in all environments (i think)

### Environment basics

Variables can point to objects that are equivalent but stored in different places in the environment.  This could be a cause of reduced performance if they are pointing to two different giant objects.

Objects that have lost their name are automatically deleted by the garbage collector.

Use `search()` to get all the parents of the global environment.  Use `ls()` to show all of the bindings in the current environment or a selected environment. Use `ls.str()` to get the structure of the environment.

```{r search, message = FALSE}
search()
ls()
ls(baseenv())
ls.str()
```

Use `rm()` to remove a binding from the environment.

```{r exists, message = FALSE}
a <- new.env()
y <- 11
exists("y", envir = a)
a$y
exists("x", envir = a, inherits = FALSE)
```


### Recursing over Environments

```{r where, message = FALSE}
require(pryr)

x <- 10
where("x")
```

### Function environments

The four function environments:

1. enclosing (`environment(f)`)
2. binding: which environment(s) have a binding to the function (`where(f)`)
3. execution: a temporary environment created during execution of the function, then thrown away; child functions can access everything in the execution environment and this environment is now their enclosing environment.
4. calling (`parent.frame()`)

Regular scoping rules only look in the function enclosing environment, not calling environment.

***

## Discussion Notes
