---
title: "Functional Programming!!!"
author: "Alathea"
date: '2014-08-20'
output: html_document
---

### *This chapter is the best!*

## Exercises

### Given a function, like `mean`, `match.fun()` lets you find a function. Given a function, can you find its name? Why doesn’t that make sense in R?

1) A function may not have a name, and
2) a function does not have a `name` attribute.

### Use `lapply()` and an anonymous function to find the coefficient of variation (the standard deviation divided by the mean) for all columns in the `mtcars` dataset.

```{r coeff_variation}
cv <- lapply(mtcars, function(x) sd(x)/mean(x))
data.frame(cv)
```

### Use `integrate()` and an anonymous function to find the area under the curve for the following functions. Use Wolfram Alpha to check your answers.

> y = x ^ 2 - x, x in [0, 10]
>
> y = sin(x) + cos(x), x in [-π, π]
>
> y = exp(x) / x, x in [10, 20]

```{r integrate}
integrate(function(x) x^2 - x, 0, 10)
integrate(function(x) sin(x) + cos(x), -pi, pi)
integrate(function(x) exp(x)/x, 10, 20)
```

### A good rule of thumb is that an anonymous function should fit on one line and shouldn’t need to use `{}`. Review your code. Where could you have used an anonymous function instead of a named function? Where should you have used a named function instead of an anonymous function?



***
***

## Discussion Notes

***
***

## Reading Notes

### Motivation

I am so happy right now. :D

The three building blocks of functional programming are:

* anonymous functions
* closures (functions written by functions)
* lists of functions

**functionals**: they take a function as an argument

### Anonymous functions

functions without a name